Omics-based Single Step Trait Prediction

### 2017-03-27 paper/manuscript.tex
First, complete draft of intro

% Introduction
\section{Introduction}

Our objectives were threefold: (1) Examine whether the single step prediction
framework can be transfered to hybrids, (2) explore the utility of single step 
predition with a quantitative predictor and (3) evaluate the impact of the
particular genetic space covered by the incomplete predictor on predictive
ability.

@@

\subsection{More detailed discussion}

\Genetics2level{Another discussion section}

>>>>

% Introduction
\section{Introduction} 
% Rationale for predictions.
Genomic prediction, pioneered by \citeNP{Meuwissen2001}, has revolutionized
animal and plant breeding by accelerating the selection of promising candidates 
without phenotypes \cite{DeLosCampos2013,Garcia-Ruiz2016} and recently emerged
as an important tool in human diagnostics \cite{DeLosCampos2010,Vazquez2016}.
Its use is largely motivated by the high costs and oftentimes long duration of 
phenotyping, which can be reduced when covering all genotypes with a predictor 
while limiting phenotyping to a few genotypes with high precision.
The performance of all other genotypes is then forecast using model parameters
obtained from thorough (repeated) $k$-fold cross-validation routines applied to 
the subset of individuals with full information.
In animal breeding, genomic selection has increased the annual selection gain 
by $\approx$ 50-100\% for yield traits and by 300-400\% for traits with low 
heritability \cite{Garcia-Ruiz2016}.
In plant breeding, the relative excellence of a breeding strategy largely 
depends on the prediction accuracy, which may vary widely among traits and 
crops \cite{Longin2015}.

% Search for better predictors
Compared to pedigree information, which represent the expected relationship
among individuals, genomic information capture Mendelian sampling and thereby 
provide an improved proxy of the realized relationship among individuals.
Nevertheless, genomic data do not exhaustively capture physiological epistasis 
\cite{Jiang2015,Guo2016,Vazquez2016}, describing the interactions 
within and between different biological strata \cite{Sackton2016}.
Such interactions were found to be pervasive throughout the genomes of yeast 
\cite{Brem2005} and humans \cite{Brown2014} and have motivated studies on the 
utility of downstream 'omics' predictors for integrating such interactions.
So far, encouraging results have been found for humans \cite{Vazquez2016}, maize 
inbred lines \cite{Guo2016} and maize hybrids \cite{Westhues2017}.

% Incomplete predictors.
Breeding programs --- particularly in animal breeding --- are characterized by
the availability of one predictor type, which is available for all recorded
genotypes while other predictors are incomplete in that they cover only a 
subset of all genotypes.
Usually, the complete predictor is pedigree information while genomic data are
the most prevalent incomplete predictor.
To utilize all available phenotypic information for model training, methods were 
needed that would allow for a combined use of genomic and pedigree information 
\cite{Hayes2009a,VanRaden2009}.

% Two-step prediction.
In the early years of genomic prediction, a two-step procedure was the most
intuitive choice for establishing a new selection scheme.
The first step constists of conventionally estimating breeding values based on
pedigree relationships and the phenotypes of relatives.
Individuals with accurate breeding values out of the first step (\textit{i.e.}
high number of offspring), enter the second step in which conventional breeding
values are regressed onto SNP genotypes.
The resulting model coefficients can then be used to predict genomic breeding 
values for solely genotyped and potentially very young individuals.
However, as the breeding values from step one enter the genomic prediction
model as phenotypes, one has to be careful about the residual error
distribution of such a model.
For instance, if the range of the prediction errors will not be identically
distributed anymore, \textit{i.e.} they should be weighted differently or
estimated individually \cite{Aguilar2010}.
One popular approach to account for this problem is deregressing the breeding
values and weighting the residuals according to the prediction error variances
\cite{Garrick2009}.

% Workings and advantages of the single step procedure.
Independently from each other, \citeNP{Legarra2009} and
\citeNP{Christensen2010} developed a single step BLUP method that blends the 
numerator relationship matrix $\mathbf{A}$ with the genomic relationship matrix
$\mathbf{G}$ in a mutual matrix $\mathbf{H}$ to use all predictor information
simultaneously.
Issues with this H-BLUP approach were the compatability of $\mathbf{A}$ with
$\mathbf{G}$ \cite{Christensen2012} and their weighting, which is not
trait-independent \cite{Ashraf2016}.
\citeNP{Fernando2014} derived the equivalent single step marker-effect-model to 
the single step breeding value model of \citeNP{Legarra2009} and 
\citeNP{Christensen2010}, which has the advantage of not requiring weights for
$\mathbf{A}$ and $\mathbf{G}$ as well as the ability to model the structure of 
the imputation error.

% -   Mention scarcity of data on single step prediction in plant populations.
Single step prediction has already become a routine procedure in animal
breeding but, to the best of our knowledge, has only been considered in plants
in a study on wheat lines \cite{Ashraf2016}. 
Our objectives were threefold: (1) Examine whether the single step prediction
framework can be transfered to plant hybrids, (2) explore the utility of single 
step predition with a quantitative predictor and (3) evaluate the impact of the
genetic space covered by the incomplete predictor on predictive ability.

@@

\Genetics2level{Superiority of single step models}

\Genetics2level{Coverage of the genetic space}
Animal populations are so large that it does not truly matter, which animals are 
being genotyped, but in plant breeding the situation is different and we would 
be interested in determining which individuals should be complemented with 
information on another predictor.

Prediction of phenotypic performance is particularly crucial in hybrid breeding
programs because the number of potential selection candidates is $n^{2}$ where
$n$ corresponds to the number of parent lines.

Given recent results on the utility of downstream 'omics' predictors in a study
on hybrid prediction in a European hybrid maize panel \cite{Westhues2017}, our 
objectives were i) to explore the benefits of a single step procedure for the 
same material and ii) to examine the influence on the composition of the set of
genotypes that is covered by more than one predictor on the success of single
step prediction.

% Special situation in hybrid breeding.
Hybrid breeding --- with noteworthy commercial applications in pig, chicken and 
maize breeding --- is a particularly challenging application field for
prediction tools.
Here, $2n$ parent individuals from two genetically distinct heterotic groups are
crossed to each other, yielding $n^{2}$ potential hybrid progeny that would
require intensive field testing.
In medium-sized plant breeding programs, the advent of the doubled haploid (DH) 
technology \cite{Wedzony2009} now allows for an annual production of 1,000 
parent lines in each heterotic group, amounting to $10^{6}$ putative hybrid
progeny.

So far, its applications in hybrid populations are limited to pig breeding
\cite{Xiang2015,Xiang2016,Tusell2016}.
% Are there applications in chicken breeding? (Hints in Christensen et al.
% 2014, GSE).
